Visual Analytics of Large Bipartite Networks assisted by Multilevel Strategies\thanks{Supported by FAPESP, project 2017/05838-3}

### 2019-05-24 paper.tex
adiantado ate soft implementation
terminado perfil geral do escrito

This paper is organized as follows: in Section~\ref{rel} the related work is examined,
while in~\ref{nom} are selected remarks about the vocabulary.
The method is delineated in Section~\ref{des} and
the software implementation is then described in Section~\ref{sof}.
Results and discussion are in~\ref{res}.
Finally, Section~\ref{con} holds concluions and further work envisioned.

@@

\noindent 
\section{Software implementation: mlvis}\label{sof}
%%%
% description of the functionalities/tools implemented,
% of the technologies and libraries used
% optimization of the computational capacity for large networks by
% means of Web GL, triangles for nodes and details on demand
We implemented the framework described in Section~\ref{des} using scientific,
database and web resources,
and made the final result available to the user within a web page
for use through simple mouse-driven actions
and requiring the instalation of no software but a web browser such as Firefox or Google Chrome.
The final result is exemplified in Figure~\ref{fpage} and next subsections
describe its functionalities and the technologies used.

@@

\subsection{The use of mlvis}
% network upload, format.
% layouts available
% algorithms available for coarsening, and their parametrization
% initialization of the visualization: click on render
% tools available and their operation

\subsection{Technologies used within mlvis}
% coarsening algorithms
% flask for server, feather for auxiliary server
% mongodb filesystem for data
% Nuxt/Vue.js for the frontend with heavy use of PIXI.js

\section{Results and discussion}\label{res}
%%%
% consistent use of overview first, details on demand
% a first contribution on ML vis. of bipartite networks
% simple interactivity strategies

% enhancements of the navigation through the use of being-developed coarsening algorithms
% the use of the interface for guiding coarsening (through markers)
% the use of the interface to analyse coarsening results

@@

\section{Conclusions and further work}\label{con}
%%%
% extension to multipartite (or heterogeneous) networks.

>>>>

This paper is organized as follows: in Section~\ref{rel} the related work is examined,
while in~\ref{nom} are selected remarks about the vocabulary.
The method is delineated in Section~\ref{des} and
the software implementation is then described in Section~\ref{sof}.
% Results and discussion are in~\ref{res}.
Finally, Section~\ref{con} holds concluions and further work envisioned.

@@

\noindent 
\section{Software implementation: mlvis}\label{sof}
%%%
% description of the functionalities/tools implemented,
% of the technologies and libraries used
% optimization of the computational capacity for large networks by
% means of Web GL, triangles for nodes and details on demand
We implemented the framework described in Section~\ref{des} using scientific,
database and web resources,
and made the final result available to the user within a web page
for use through simple mouse-driven actions
and requiring the instalation of no software but a web browser such as Firefox or Google Chrome.
The final result is exemplified in Figures~\ref{fpage0} and~\ref{fpage} and next subsections
describe its functionalities and the technologies used.

\begin{figure}[!h]\centering
 \includegraphics[width=\textwidth]{fpage0}
  \caption{The navigation pathway of the bipartite network obtained through a multilevel strategy
  and auxiliary tools.
  }\label{fpage0}
\end{figure}

@@

\subsection{The use of mlvis}
% network upload, format.
% layouts available
% algorithms available for coarsening, and their parametrization
% initialization of the visualization: click on render
% tools available and their operation
The exposition on this section is suported by Figure~\ref{usage}.
The analyst first selects a network of interest to be explored.
One may also upload a network and then select it on the same drop-down menu.
The software is then ready to render the network on the screen for arbitrary
navigation, but the user may wish to select one of the available network layout
algorithms and tune the coarsening to be performed.
Upon request given by hiting the ``render network'' button,
the network is mapped to the screen.
The subsequent usage relies on the manipulation of the visual mapping presented on the
canvas by means of actions
performed through managing the tools and related mouse operations.

There are multiple levels of representation, which is entailed by the multilevel strategy.
This makes pertinent that actions be performed in only one of the available levels.
Thus, there is always a ``level on focus'' in which the tools perform actions, such
as link coloring or opening supernodes to expose their child nodes.
The first button on the toolbar is thus for changing the level on focus.
The second set of buttons are for changing how the nodes are visualized:
color, size, if size is related to degree or number of children, restore node size,
node transparency, and rotation.
The third set of tool buttons if for links: color, thickness, transparency, if thickness and
transparency are related to link weight, and restor link properties.

The fourth set of tool buttons is more specific to the needs of the MlBiNetViz, and
are selected to make mouse operations entail special processes and transformations of the visualization.
The first button is selected for the interface to return core information from a clicked node:
id, number of children, id of parent, number of neighbors (degree), sum of weight of links (strength), and other information the analysts report necessity of.
The second tool button is seleted to move nodes: specific nodes are individually moved if clicked and dragged; a rectangular region is made available for moving all nodes therein if an empty region of the canvas is clicked and the mouse cursor dragged before releasing the button.
The third tool button is selected for opening a supernode and exposing their child nodes and links among them:
a region is defined by the user by clicking on an empty space on the canvas and gragging the mouse before releasing, the supernodes therein are joined into a rectangular region that contains all of them, and their child nodes (on a finer level of detail) are drawn inside such rectangular region.
Open supernodes may be joined using the same tool, but it was found convenient to make available a
fourth button to assist in this task: if this tool is selected, one may join two open supernodes by
clicking on them in sequence.
The last tool button is selected to perform three operations on open supernodes:
if an open supernode is clicked on and dragged, the rectangular shape is resized accordingly;
if there is no drag, i.e. the super node is clicked on and released, the links from it to the other
supernodes are attached to the next counterclockwise corner of its rectagular region.
When the open supernode is resized, the layout of its child nodes is optimized for the drawn region.
This is most useful when used in combination with the drag tool: on the finer level of detail, the nodes are repositioned into a more compact disposition. Then upon a (potentially minimal) supernode
risize, the child nodes are made to fit the whole rectangular area.

The last set of tools acts on the whole canvas: they are dedicated to zoom and left-right, up-down pan, changing the color of the background, and toggling between current zoom and pan status and the initial setting.

There are many subleties on the orchestrated use of these tools.
As a facilitated means to convey the navigation possibilities and
procedures available, the user is invited to watch a demo video in~\cite{yvideo},
and use the MlBiNetViz in the instance we made availabel~\cite{mbnvpage}.
The more seasoned software developer may browser the code, upload a local instance
of the software, and make changes to its source code~\cite{mbnvcode}.

\subsection{Software technologies employed within mlvis}
% coarsening algorithms
% flask for server, feather for auxiliary server
% mongodb filesystem for data
% Nuxt/Vue.js for the frontend with heavy use of PIXI.js
Most important parts of MlBiNetViz are written in a combination of JavaScript and Python:
Vue.js (set up by Nuxt.js) is used to assist in the client (frontend),
the server (backend) is a Flask Python server, used to perform
specific or heavy calculations.
There is a secondery server, a node FeatherJS, used to facilitate contact with the database
and real-time multi-user interaction.
The data is stored in a MongoDB database and ordinarily in the filesystem,
while the multi-user interaction was deactivated for now to avoid unecessary complexity.
The coarsening algorithms for bipartite networks are available in~\cite{bialgs}
and are used by the Flask server.
The fast WebGL 2D rendering on the canvas is performed using Pixi.js.
This is a very brief account of the styles and software resources used to implement
MlBiNetViz. The interested reader is welcome to browse the software code and documentation
and contact the authors~\cite{mbvncode}.

% \section{Results and discussion}\label{res}
%%%
% consistent use of overview first, details on demand
% a first contribution on ML vis. of bipartite networks
% simple interactivity strategies
% enhancements of the navigation through the use of being-developed coarsening algorithms
% the use of the interface for guiding coarsening (through markers)
% the use of the interface to analyse coarsening results

@@

\section{Conclusions and further work}\label{con}
%%%
% extension to multipartite (or heterogeneous) networks.
The resulting conceptual organization is very consistent with the Shneiderman's mantra,
as exemplified the by the implemented software.
Also, it is the first contribution to bipartite networks visual analytics
assited by multilevel strategies.
The interface developed is efficiently managed through simple interactive operations.

The navigation may be enhanced by the use of in-development coarsening algorithms.
The visualization pathway should be used to assist multilevel strategies by enabling the user
to select nodes as pivots to guide the coarsening procedure.
Also, the interface is envioned to assist the analysis of coarsening results.
The bipartite framework is naturally extended to heterogeneous (or multipartite) networks,
which may be a research route we will follow if there is enough interest of the researchers involved.
